Fail Faster

### 2025-03-24 paper/paper.tex
13x

%\jwc{``Property based testing is a race against time...''}
Property-based testing (PBT) is a software testing technique that uses
random test inputs to validate logical specifications. Recent studies on PBT usage in industry~\cite{inpractice} have shown that
practitioners often interleave PBT with their software development process, sometimes
testing their properties every time they save their code. This means that PBT is often given a very short time
window to find bugs---e.g., between 50 milliseconds and 30 seconds~\cite{inpractice}.
The faster a test fails, the better.
% The goal of PBT is to fail tests as fast as possible. \jwc{First cut at ``fail fast''}
\iflater 
\bcp{Feels like we're missing a trick in the transition here.  Can we
  say, ``Because performance is important, PBT libraries are already
  quite tuned.  [Can we substantiate that??]  But more is
  possible...'' }
\fi

@@

This is due to two major sources of inefficiency.
First, the same high-level design approach that gives generator libraries their
flexibility also introduces layered abstractions and indirections that are difficult for
compilers to optimize\proposecut{, shifting work to run time}.
Second, the \rand that a generator uses to make choices can
constitute an large proportion of its running time, magnifying the impact of an inefficient one. \jwc{Object of ``one'' unclear, intended to be ``\rand{}''.}
\iflater
\hg{TODO later, I'm not 100\% sure it's clear how this particular question is
raised by the above. We could strengthen that connection}
\fi

@@

In summary, we show that PBT generator DSLs incur significant performance costs across languages, and present a general technique for improving their efficiency while preserving their idiomatic abstractions. \jwc{needs a bit about RNG --- BCP thoughts?}
We make the following contributions:
\begin{enumerate}
    \item We identify two key sources of inefficiency in PBT generator libraries, which can significantly impact performance:
    abstraction overhead and \rand choice.
    \item We present \name{}, a technique for eliminates the abstraction overhead of generators. We implement \name{} in both OCaml and Scala 3, showcasing its cross-language applicability.
    \item We demonstrate that writing generators using \name{} and fast \rands{} yields significant performance improvements.
    We precisely quantify the impact of each optimization separately, and explain how they play distinct yet complementary roles in improving generator performance.
    Finally, we show that these performance improvements extend to improved bug-finding speed, finding some bugs as much as \todo{\%\%} faster.
\end{enumerate}

>>>>

%\jwc{``Property based testing is a race against time...''}
Property-based testing (PBT) is a software testing technique that uses
random test inputs to validate logical specifications. Recent studies on PBT usage in industry~\cite{inpractice} have shown that
practitioners often interleave PBT with their software development process, sometimes
testing their properties every time they save their code. This means that PBT is often given a very short time
window to find bugs---e.g., between 50 milliseconds and 30 seconds~\cite{inpractice}.
The faster a test fails, the better.
% The goal of PBT is to fail tests as fast as possible. \jwc{First cut at ``fail fast''}
% \iflater 
% \bcp{Feels like we're missing a trick in the transition here.  Can we
%   say, ``Because performance is important, PBT libraries are already
%   quite tuned.  [Can we substantiate that??]  But more is
%   possible...'' }
% \fi

@@

This is due to two major sources of inefficiency.
First, the same high-level design approach that gives generator libraries their
flexibility also introduces layered abstractions and indirections that are difficult for
compilers to optimize.
Second, the calls a generator makes to its \rand{} can constitute a large 
proportion of its running time, magnifying the cost of an inefficient 
implementation.

@@

In summary, we show that PBT generator DSLs incur significant performance costs across languages, and present a general technique for improving their efficiency while preserving their idiomatic abstractions. \jwc{needs a bit about RNG --- BCP thoughts?}
We make the following contributions:
\begin{enumerate}
    \item We identify two key sources of inefficiency in PBT generator libraries, which can significantly impact performance:
    abstraction overhead and \rand choice.
    \item We present \name{}, a technique for eliminates the abstraction overhead of generators. We implement \name{} in both OCaml and Scala 3, showcasing its cross-language applicability.
    \item We demonstrate that writing generators using \name{} and fast \rands{} yields significant performance improvements.
    We precisely quantify the impact of each optimization separately, and explain how they play distinct yet complementary roles in improving generator performance.
    Finally, we show that these performance improvements extend to improved bug-finding speed, finding some bugs as much as $13\times$ faster.
\end{enumerate}
